Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-10-16 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update advection stats to reflect new advection fix (277963c)

The initial thermodynamic conditions have a temperature at $z=0$ of $\theta_0 = \SI{288}{\kelvin}$ and constant static stability with $N = \SI{0.01}{\per\second}$ everywhere, such that
\begin{align}
	\theta(z) = \theta_0 \exp \left( \frac{N^2}{g} z \right) \label{eqn:thermal-profile}
\end{align}
for $z > 0$.
Potential temperature values are prescribed at the inlet and upper boundary using equation~\ref{eqn:thermal-profile}, and a zero gradient boundary condition is applied at the outlet.  At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~\ref{eqn:thermal-profile}.
For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.

@@

% TABLES
\begin{table*}
	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                  & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.107   & 0.0146          & 0.00784        & 0.00784     \\
                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & {unstable}     & {---}        \\
	          &                     & min              & 0        & -0.0245 & -0.120          & {unstable}     & {---}        \\
		  &                     & max              & 1        & 0.985   & 0.950           & {unstable}     & {---}        \\
		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
\hline
\end{tabular}
\end{table*}

>>>>

The initial thermodynamic conditions have constant static stability with $N = \SI{0.01}{\per\second}$ everywhere, such that
\begin{align}
	\theta(z) = \theta_0 \exp \left( \frac{N^2}{g} z \right) \label{eqn:thermal-profile}
\end{align}
where the temperature at $z=0$ is $\theta_0 = \SI{288}{\kelvin}$.
Potential temperature values are prescribed at the inlet and upper boundary using equation~\ref{eqn:thermal-profile}, and a zero gradient boundary condition is applied at the outlet.  At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~\ref{eqn:thermal-profile}.
For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.

@@

% TABLES
\begin{table}
	\caption{\TODO{mesh statistics.  BTF max/min cell area ratio is very close to 1 irrespective of resolution.}}
	\label{tab:gw-meshes}
%
\centering
\begin{tabular}{S S S S}
\hline\hline

{$\Delta z$ (\si{\meter})} & {cut cell max/min cell area ratio} & {BTF nonortho} & {cut cell nonortho} \\
\hline
300	& 4.325 & 10.2	& 7.3 \\
250	& 5.422 & 10.8	& 20.9 \\
200	& 25.36 & 10.8	& 31.9 \\
150	& 10.03 & 11.0	& 29.4 \\
125	& 10.11 & 11.1	& 24.37 \\
100	& 36.27 & 11.1	& 32.2 \\
75	& 52.89 & 11.2	& 34.2 \\
50	& 75.25 & 11.2	& 31.7 \\
\hline
\end{tabular}
\end{table}

\begin{table*}
	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                  & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
                  &                     & min              & 0        & -0.0464 & -0.0106         & -0.000674      & -0.00674    \\
                  &                     & max              & 1        & 0.922   & 0.982           & 0.983          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & {unstable}     & {---}        \\
	          &                     & min              & 0        & -0.0245 & -0.120          & {unstable}     & {---}        \\
		  &                     & max              & 1        & 0.985   & 0.950           & {unstable}     & {---}        \\
		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
\hline
\end{tabular}
\end{table*}
